### 2018-12-12 paper/main.tex

%!TEX options = --shell-escape

\documentclass[bachelor]{thesis-uestc}
%----文内引用符号
\usepackage[utf8]{inputenc}
\usepackage{cleveref}
\crefname{section}{§}{§§}
%----------

\title{密文重复数据删除机制的频率攻击方法}
\author{任彦璟}
\advisor{李经纬\chinesespace 副教授}
\school{计算机科学与工程学院(网络空间安全学院)}
\major{信息安全}
\studentid{2015040101018}

\begin{document}

\makecover

\begin{chineseabstract}

加密重复数据删除旨在解决大规模数据存储系统中的安全性和存储效率：它确保每个明文都由明文本身内容派生的对称密钥加密，从而为数据存储提供机密性保证，同时保留重复数据删除的存储节省效率。但是，加密重复数据删除的确定性特性也会泄漏明文的频率，从而使加密重复数据删除容易受到频率分析的影响。在本文中，我们重新审视了频率分析导致的加密重复数据删除的安全漏洞。我们认为加密重复数据删除可能更容易受到精心设计的频率分析攻击，因此攻击提供了高可信度，可确保每个推断的明文确实对应于目标密文的概率很高（即从统计角度来看，高精度） 。为此，我们提出了两种新的频率分析攻击，它们可以适应实际重复数据删除工作负载的特性，从而提高频率分析的准确性。我们根据实际情况评估针对实际存储工作负载的建议攻击，并提供有关如何带来实际损失的观察。

\chinesekeyword{加密重复数据删除, 频率分析攻击, 聚类分析}
\end{chineseabstract}

\begin{englishabstract}
Encrypted deduplication aims to address security and storage efficiency in large-scale data storage systems: it ensures that each plaintext is encrypted by a symmetric key that is derived from the content of the plaintext itself, so as to provide confidentiality guarantees for data storage while preserving the storage saving effectiveness of deduplication. However, the deterministic nature of encrypted deduplication also leaks the frequencies of plaintexts, thereby making encrypted deduplication vulnerable to frequency analysis. In this paper, we revisit the security vulnerability of encrypted deduplication due to frequency analysis. We argue that encrypted deduplication can be even more vulnerable to carefully crafted frequency analysis attacks, such that the attacks provide high confidence of ensuring that each inferred plaintext indeed corresponds to the target ciphertext with a high probability (i.e., high precision from a statistical perspective). To this end, we propose two new frequency analysis attacks that adapt the characteristics of practical deduplication workloads to increasing the severity of frequency analysis. We empirically evaluate our proposed attacks against real-world storage workloads, and provide observations on how they bring actual damages.

\englishkeyword{Encrypted deduplication, Frequency analysis attacks, Cluster analysis}
\end{englishabstract}

\thesistableofcontents

\thesischapterexordium

\input{introduction}

%\thesisacknowledgement

\nocite{*}
\bibliographystyle{thesis-uestc}
\bibliography{reference}

\thesisappendix

%\thesisloadachievement{publications}

%\thesistranslationoriginal
%\section{}
%
%
%\thesistranslationchinese
%
%\section{}

\end{document}